Supplemental Information for: The Magnetorotational Instability Prefers Three Dimensions

### 2019-08-21 paper/supplemental.tex
tidy up supp. material

\section{Asymptotic Calculation}
\label{sec:asymp}
Define the domain as $ - d/2 < x < d/2$. Make it symmetric so to make the shear simpler. The shear has no net. Otherwise it means we are not exactly in the rotating frame we thought we were. 
\Beq
\int_{-d/2}^{d/2} S x \text{d} x  \ = \ 0
\Eeq
This is very important later.

The full linear equations:
\Beq
\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B \pd{z} b_{x} &=& 0\\
\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B \pd{z} b_{y} &=& 0\\
\pd{t} w + S x \pd{y} w + \pd{z} p - B \pd{z} b_{z} &=& 0 \\
\pd{x} u + \pd{y} v + \pd{z} w  &=& 0\\
\pd{t} b_{x} + S x \pd{y} b_{x} - B \pd{z} u &=& 0\\
\pd{t} b_{y} + S x \pd{y} b_{y} - B \pd{z} v - S b_{x}  &=& 0\\
\pd{t} b_{z} + S x \pd{y} b_{z} - B \pd{z} w &=& 0
\Eeq

Assume all variable take the form
\Beq
g = \hat{G}(x) e^{ i ( \omega t + k z + \ell y ) } 
\Eeq

Define the ``parameter''
\Beq
\omega_{S} \ \equiv \ \omega + S x \ell 
\Eeq

We can use equations (2)--(7) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
\Beq
\hat{V} &=& \frac{i \left(k^2 \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\ell  \omega _S \hat{U}'(x) \left(B^2 k^2-\omega
   _S^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
   \left(B^2 k^2-\omega _S^2\right)}\\
\hat{W} &=& -\frac{i k \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S
   \left(B^2 k^2-\omega _S^2\right)}\\
   \hat{P} &=& -\frac{i \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S) \omega
   _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2}\\
   \hat{B}_{x} &=& \frac{B k \hat{U}(x)}{\omega _S}\\
   \hat{B}_{y} &=& -\frac{i B k \left(\hat{U}(x) \left(B^2 k^2 S \ell ^2+\omega _S^2
   \left(f k^2-S \ell ^2\right)\right)+\ell  \omega _S
   U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
   \left(B^2 k^2-\omega _S^2\right)}\\
   \hat{B}_{z} &=& -\frac{i B k^2 \left(\ell  \hat{U}(x) \left(B^2 k^2 S-(f+S)
   \omega _S^2\right)+\omega _S U'(x) \left(\omega _S^2-B^2
   k^2\right)\right)}{\left(k^2+\ell ^2\right) \omega _S^2
   \left(B^2 k^2-\omega _S^2\right)}
\Eeq

We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}$.

When we do this, we get something of the form 
\Beq
\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat(U)(x) \ = \ 0
\Eeq
where
\Beq
\alpha(x) &=& \frac{i \left(B^2 k^2-\omega _s^2\right)}{\left(k^2+\ell
   ^2\right) \omega _s} \\ 
   \beta(x) &=& -\frac{2 i B^2 k^2 S \ell }{\left(k^2+\ell ^2\right) \omega
   _s^2} \\
   \gamma(x) &=& \frac{i \left(\omega _s^4 \left(\frac{f^2
   k^2}{\left(k^2+\ell ^2\right) \left(B^2 k^2-\omega
   _s^2\right)}+1\right)-k^2 \omega _s^2 \left(B^2+\frac{f
   S}{k^2+\ell ^2}\right)+\frac{2 B^2 k^2 S^2 \ell
   ^2}{k^2+\ell ^2}\right)}{\omega _s^3}
\Eeq

We can eliminate the first-order term via
\Beq
\hat{U}(x) \ = \ \chi(x) \psi(x) 
\Eeq
if 
\Beq
\chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
\psi''(x) + \Phi(x) \psi(x) \ = \ 0
\Eeq
\Beq
\Phi(x)  \ = \ -\frac{-2 \beta (x) \alpha '(x)+2 \alpha (x) \left(\beta'(x)-2 \gamma (x)\right)+\beta (x)^2}{4 \alpha (x)^2}
\Eeq
Putting everything together 
\Beq
\Phi(x) \ = \ \tfrac{k^2 \omega _S^2 \left(2 B^2 \left(k^2+\ell
   ^2\right)+f (f+S)\right)-B^2 k^2 \left(B^2 k^2
   \left(k^2+\ell ^2\right)+S \left(f k^2-S \ell
   ^2\right)\right)-\left(k^2+\ell ^2\right) \omega
   _S^4}{\left(\omega _S^2-B^2 k^2\right){}^2}
\Eeq

Now we want to rescale everything in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B^2}{d^2 f} ( 1 + \eps^{2} R) 
\Eeq 
\Beq
\omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell 
\Eeq
Also 
\Beq
q  \ \equiv  \ \frac{\pi ^2 B^2}{d^2 f^{2}}
\Eeq

Expanding to $\mathcal{O}(\eps^{2})$, 
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B^4 k^4+B^2 f^2 q \left(k^2 R+q
   \ell ^2\right)+f^2 (q+1) (\omega -f q x \ell
   )^2\right)}{B^4 k^2}
\Eeq

Now do the perturbations

\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 
\Eeq
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0 
\Eeq

Remember we are on $-d/2 < x < d/2$ Therefore 
\Beq
\psi_{0}(x) \ = \ \cos(\pi x / d)
\Eeq
This satisfies the equation and boundary condition. The solvability condition for $\psi_{2}(x)$ is

\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq

This works out to 
\Beq
\left(\pi ^2 B^2+d^2 f^2 \right) \omega^{2} - B^4 k^2 \left(d^2 k^2-\pi ^2 R\right)   \ + \ \frac{\pi ^2 B^4 \ell ^2}{12}  \left(\left(6+\pi ^2\right) q+\pi ^2-6\right) \ = \ 0
\Eeq

\Beq
\omega^{2} \ = \ \frac{q\,B^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
\Eeq

>>>>

\section{Asymptotic Calculation}
\label{sec:asymp}
Define the domain as $ - d/2 < x < d/2$. Being in the rotating frame means the shear has no net
\Beq
\int_{-d/2}^{d/2} V_{y}(x) \text{d} x  \ = \ 0, \quad \implies  \quad V_{y}(x) \ = \ S\, x
\Eeq
The full linear ideal equations are:
\Beq
\pd{t} u + S x \pd{y} u - f v + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
\pd{t} v + S x \pd{y} v + (S+f) u + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\\
\pd{t} w + S x \pd{y} w + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
\pd{x} u + \pd{y} v + \pd{z} w  &=& 0\\
\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} u &=& 0\\
\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v - S b_{x}  &=& 0\\
\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} w &=& 0
\Eeq

Assume all variable take the form
\Beq
g = \hat{G}(x) e^{ i ( \omega t + k z + \ell y ) } 
\Eeq
Define the frequency ``parameters'' 
\Beq
\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k
\Eeq
We can use equations (2)--(7) to find all amplitudes in terms of $\hat{U}(x)$ and $\hat{U}'(x)$. 
\Beq
\hat{V}(x) &=& \frac{i \left(\ell 
   \hat{U}'(x) + k^{2} \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)} \hat{U}(x)\right)}{k^2+\ell ^2} \\
\hat{W}(x) &=& \frac{i k \left(\hat{U}'(x)- \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)}  \ell 
   \hat{U}(x)\right)}{k^2+\ell ^2} \\
   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
   \left(\hat{U}'(x) -  \ell \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)}
   \hat{U}(x)\right)}{\left(k^2+\ell ^2\right)
   \omega _S} \\
   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{U}(x)}{\omega _S} \\
   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}(x)}{\omega _S} -\frac{i S \omega _A \hat{U}(x)}{\omega _S^2} \\
   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{W}(x)}{\omega _S}
   \Eeq
We can then substitute everything into equation (1) to get a second-order equation for $\hat{U}(x)$ of the general form 
\Beq
\alpha(x) \hat{U}''(x) + \beta(x) \hat{U}'(x) + \gamma(x) \hat{U}(x) \ = \ 0
\Eeq
where
\Beq
\alpha(x) &=& \omega _S^2-\omega _A^2 \\ 
   \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k^2}{\omega
   _S^2-\omega _A^2}-\frac{2 S^2 \ell
   ^2}{\omega _S^2}+k^2+\ell ^2\right)+f k^2
   (f+S)-\left(k^2+\ell ^2\right) \omega _S^2
\Eeq

We can eliminate the first-order term via
\Beq
\hat{U}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0
\Eeq
\Beq
\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
   \left(\beta '(x)-2 \gamma (x)\right)+\beta
   (x)^2}{4 \alpha (x)^2}
\Eeq
Putting everything together 
\Beq
\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
   _S^2}{\left(\omega _A^2-\omega
   _S^2\right){}^2} + k^{2} + \ell^{2}
\Eeq

Rescale everything in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R) 
\Eeq 
\Beq
\omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell 
\Eeq
Also 
\Beq
q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}
\Eeq
Expanding to $\mathcal{O}(\eps^{2})$, 
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \frac{\epsilon ^2 \left(-B_{0}^4 k^4+B_{0}^2 f^2 q \left(k^2 R+q
   \ell ^2\right)+f^2 (q+1) (\omega -f q x \ell
   )^2\right)}{B_{0}^4 k^2}
\Eeq
Going order-by-order
\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0 
\Eeq
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0 
\Eeq
Therefore 
\Beq
\psi_{0}(x) \ = \ \cos(\pi x / d)
\Eeq
This satisfies the equation and boundary condition. The solvability condition for $\psi_{2}(x)$ is 
\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq
This implies 
\Beq
\omega^{2} \ = \ \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left(\frac{d^2}{\pi^{2}} k_z^2-R\right) \ - \ \frac{\left(6+\pi ^2\right) q+\pi
   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
\Eeq
